Supplemental data to ``\emph{k}-mer-based GWAS enhances causal variant and candidate gene discovery in soybean''

### 2023-03-03 additional_files/main_text.tex
Make a few changes in the text and figures upon proofreading the manuscript

Genome-wide association studies (GWAS) are powerful statistical methods
that detect associations between genotype and phenotype at genome scale. Despite
their power, GWAS frequently fail to pinpoint the causal variant or the gene
controlling a trait at a given locus in crop species. Assessing genetic variants beyond
single-nucleotide polymorphisms (SNPs) may alleviate this problem,
for example by including structural variants (SVs) that may have a greater functional impact.
In this study, we tested the potential
of SV- and \emph{k}-mer-based GWAS in soybean by applying these methods to 13
traits on which SNP-based GWAS had been performed previously.
We assessed the performance of each GWAS approach based on results at loci for which
the causal genes or variants were known from previous genetic studies. We found
that \emph{k}-mer-based GWAS was the most versatile approach and the best at pinpointing
causal variants or candidate genes based on the most significantly
associated \emph{k}-mers. Given the traits that we analyzed, SV-based GWAS did not add any value
compared to \textit{k}-mer-based GWAS and may not generally be worth the investment in time and
computational resources required to discover and genotype such variants.
Moreover, \textit{k}-mer-based analyses
identified promising candidate genes for loci related
to pod color, pubescence form, and resistance to the oomycete
\textit{Phytophthora sojae}. Despite promising results,
significant challenges remain regarding the downstream analysis of
\emph{k}-mer-based GWAS, notably in associating significant \emph{k}-mers with
sequence variation and ensuring that all variation types are adequately
detected from \emph{k}-mer occurrence patterns. Together, our results suggest
that coupling \emph{k}-mer- and SNP/indel-based GWAS may represent a
powerful approach for discovering candidate genes in crop species.

@@

The inclusion of structural variants (SVs) as genotype data in GWAS may partly overcome
this issue by providing a set of variants with a potentially high functional impact.
SVs include any variant involving a difference of at least 50 nucleotides, such as
deletions, insertions, or inversions \citep{ho2020}. Such variants are known to
have large phenotypic impacts by disrupting coding or regulatory sequences
\citep{marroni2014}, and have been identified as causal variants in traits such
as resistance to soybean cyst nematode \citep{cook2012}, aluminum tolerance in
wheat \citep{maron2013}, and branching in maize \citep{studer2011}.  SVs have
already been used in GWAS in studies with promising results
\citep[e.g.][]{zhang2015, akakpo2020, dominguez2020, liu2020-pangenome}, but
their inclusion in GWAS is the exception rather than the norm.

Despite their phenotypic impact, SVs are still largely understudied compared to SNPs
because of the difficulty of discovering and genotyping them with
accuracy. Comprehensive detection of SVs requires a combination of several
methods that often need to be tailored to the study species and data available
\citep{alkan2011, ho2020}. Even then, assessments of the performance of SV
discovery and genotyping generally reveal subpar sensitivity and precision
\citep[e.g.][]{cameron2019, chaisson2019, kosugi2019}. Moreover, despite recent
improvements in computational approaches for genotyping SVs
\citep[e.g.][]{siren2021, ebler2022}, such approaches remain relatively inaccurate
compared to SNP genotyping workflows.

Alternative approaches that allow one to assess a large spectrum
of variants in GWAS without genotyping SVs at population scale could be useful.
\cite{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
represent such an alternative method. This approach relies on the presence or
absence of \emph{k}-mers observed in sequence reads instead of genotypes for use in
GWAS. Once a presence/absence table of \emph{k}-mers has been generated for the
population under study, this table can be used in association analyses
similarly to SNP or SV genotype calls. One advantage of using \emph{k}-mers is
that they can act as a molecular signature for any type of variant as long as
these variants result in presence/absence of \emph{k}-mers of a given length.
Another advantage is that \emph{k}-mers are not tied to a specific genomic
location and can thus be used to query associations between genotype and
phenotype at genomic locations that are not found in a reference genome.  A
few studies have applied this \emph{k}-mer-based GWAS approach
\citep[e.g.][]{tripodi2021, colque2021} and other similar approaches have been
developed \citep{rahman2018, he2021}, but similarly to the use of SVs in GWAS,
such approaches have not been widely adopted yet.

In this study, we assessed the potential of SV- and \emph{k}-mer-based GWAS in
a major crop, soybean (\emph{Glycine max}).  In order to test these methods, we
analyzed ten qualitative traits \citep{bandillo2017} and two quantitative
traits \citep{bandillo2015} that have been previously studied using a SNP array
developed for soybean \citep{song2013}. In addition, we analyzed a quantitative trait
(horizontal resistance to \emph{Phytophthora sojae}) that has been previously
studied using whole-genome sequencing (WGS) data \citep{deronne2022}.
Since the underlying genes and causal variants are already known for several of
these traits, they provide a good test case for the performance of GWAS methods.
The objectives of our study were to:

@@

We analyzed a total of 13 traits in a population of 363 \textit{G. max}
accessions using three GWAS approaches that differed in the genotypes used. The
first approach used SNP and indel genotypes called by Platypus. Despite both
SNPs and indels being included in this approach, we will refer to this approach
as SNP-based for simplicity. Our SV-based workflow used various sources of data
for SV discovery and a single program (Paragraph) for genotyping SVs at
population scale. Finally, the third GWAS approach relied on the
presence/absence of \textit{k}-mers as genotypes.  Because of the large number
of \textit{k}-mers analyzed, we only report results (e.g. in Manhattan plots)
for those that were considered significantly associated with traits of
interest.

@@

% Pubescence color - T locus
\cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
\textit{T} locus for pubescence color and found a 1-bp deletion as a putative
causal variant at this locus. For all three GWAS approaches, we found a signal
overlapping this gene (Additional file 1: Figure \ref{pubescence_color_all_manhattan}). In
particular, the regions defined by the 1\% most associated variants
identified by the SNP- and \textit{k}-mer-based approaches included
Glyma.06g202300 (Additional file 1: Figure \ref{pubescence_color_all_T_signal}). While the most
significant \textit{k}-mers did occur within the sequence of this gene, they
did not correspond to the causal variant (Additional file 1: Figure
\ref{pubescence_color_all_T_gene}d).  Still, the fourth and fifth most strongly
associated \textit{k}-mers corresponded to the causal variant (Additional file 1: Figures
\ref{pubescence_color_all_T_gene}d and \ref{pubescence_color_all_T_kmers}). We
observed that not all accessions with gray pubescence harbored the deleted
nucleotide (Additional file 1: Figure \ref{pubescence_color_all_T_kmers}c), which could indicate
either that other causal variants are resulting in a non-functional F3’H in
this population, or that the causal variant lies elsewhere. As for SNP-based GWAS,
the most significant variant within the sequence of Glyma.06g202300 was the
causal variant, however it was not the most significantly associated variant
overall (Additional file 1: Figures \ref{pubescence_color_all_T_signal}b and
\ref{pubescence_color_all_T_gene}b). SV-based GWAS did not detect the causal variant
at this locus, as expected from its size.

@@

% Pubescence color - Td locus
We conducted a separate GWAS analysis contrasting only accessions with tawny
and light tawny pubescence because the \textit{Td} locus is known to control
these differences in color. Accordingly, we detected signals in this region
using all three approaches, but only SNP- and \textit{k}-mer-based GWAS
identified a signal that overlapped Glyma.03g258700 (Figure
\ref{pubescence-color-main-figure}a-b, and Additional file 1: Figures \ref{pubescence_color_nogray_manhattan} and
\ref{pubescence_color_nogray_Td_signal}).  While the most significant
\textit{k}-mer identified the causal variant documented at this locus
\citep{yan2020}, it was the fourth most significantly associated variant for
the SNP-based approach (Figure \ref{pubescence-color-main-figure}b-d,Additional file 1: Figure 
\ref{pubescence_color_nogray_Td_gene}).  Interestingly, some of the accessions
bearing the haplotype associated with tawny pubescence displayed light tawny
pubescence, suggesting that there may be more than one causal variant at this
locus (Figure \ref{pubescence-color-main-figure}e).

@@

The difference in size between the CNV reported by \cite{liu2020ps} and our
results may be due to the presence of some unresolved sequence (a stretch of
4,635 'N' nucleotides) over that interval in the reference sequence that we used.
Because our SV filtering process removed any SV overlapping 'N' nucleotides in
the reference, we reasoned that the CNV may have been found by our pipeline but
filtered out later on. Indeed, found a 42.8-kb duplication corresponding
to the one described above among the genotype calls made by smoove. An association
analysis of that single variant using GAPIT found a \textit{p}-value of 1.9e-23,
which would have made it the most significant SV. In this particular case, the
failure of the SV-based GWAS analysis to find identify the causal variant was due
to our filtering parameters and not to limitations of the SV genotype calls per se.

@@

\begin{figure}
	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
	\textit{Pa1} locus for pubescence form. (a) Genome-wide Manhattan plots.
	Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{Pa1} and \textit{Pa2} loci. (b) Zoomed-in Manhattan plots at the \textit{Pa1} locus.
	The vertical dotted lines indicate the location of the
	Glyma.12g213900 gene suggested as a candidate for this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of SNPs, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.12g213900. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrows indicate the
	direction of transcription.
	(d) Manhattan plots at the location of the Glyma.12g213900 gene.
	Vertical lines indicate the location of two non-synonymous
	SNPs associated with appressed pubescence, which may represent causal variants
	at this locus. Panels (c) and (d) share the same axis.}
\end{figure}

@@

In conclusion, we used SV-based and \emph{k}-mer-based GWAS to study ten
qualitative traits and three quantitative traits in soybean cultivars and
compared the results to traditional SNP-based GWAS. While \emph{k}-mer-based
GWAS proved to be a powerful approach in pinpointing the causal variation or
genes associated with known loci, it is unclear whether conducting SV-based
GWAS is worth the investment. In addition, we used the results from the
\emph{k}-mer- and SNP-based GWAS to suggest candidate genes for a few classical
loci that have yet to be cloned. Based on our results, we believe that an
optimal workflow may involve conducting SNP-based and \emph{k}-mer-based GWAS
in parallel to identify significant signals and candidate genes, potentially in
combination with SV datasets. Given the large number of samples for which WGS
data are now available in several species, applying \emph{k}-mer-based GWAS to
leverage already existing sequencing and phenotypic data appears feasible and
promising.  As part of our work, we developed several computational tools that
should help other researchers with the downstream analysis of
\emph{k}-mer-based GWAS. However, much work remains to be done in
developing state-of-the-art methods for the downstream analysis of significant
\emph{k}-mers. In particular, better approaches are needed for linking
\emph{k}-mers to sequence variation and moving from presence/absence-based
methods to count-based methods.

@@

The SVs called using each of these approaches were merged using SVmerge
\citep{wong2010} and the resulting set of candidate SVs was used for
genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
from the mapped Illumina reads for all accessions. Detailed methods regarding
the SV discovery and genotyping steps can be found in Additional file 1.

@@

\begin{enumerate}
	\item \texttt{FILTER} field set to \texttt{PASS}
	\item minor allele frequency (MAF) $\geq$ 0.02
	\item fraction of missing genotypes $\leq$ 0.5
	\item heterozygosity rate $\leq$ 0.2.
\end{enumerate}

@@

Because of clear discrepancies between the observed haplotypes and the reported
phenotypes for some simple traits such as flower color and seed coat color, we
suspected possible errors in the identity of some of the accessions. To
investigate this, we compared the SNP genotypes derived from the SoySNP50K
array and our WGS data at over 32,000 SNPs, and identified 24 accessions with <
90\% concordance between the two genotype datasets (see Additional file 1 for
detailed methods). These samples were removed from the dataset in addition to another
accession that exhibited an atypical GC content suggesting contamination.
Following these filtering steps and the removal of HN019, 363 accessions
remained for GWAS analyses. The concordance between WGS data and SoySNP50K
genotypes is included in Additional file 3 and averaged 98.3\% among retained
lines. The phenotypic data used in this study can be found in Additional file 4.

@@

\emph{k}-mers are not intrinsically associated with a particular genomic region
and a tailored analysis was therefore required to associate them with genomic
coordinates. To do so, we queried the reads (in BAM format) of all accessions for matches
to significant \emph{k}-mers for a given trait. Because this is a
computationally demanding operation (there were typically tens of thousands of
significant \emph{k}-mers per trait), we developed a C program called katcher
(\url{https://github.com/malemay/katcher}) that
uses the htslib library \citep{bonfield2021} for efficient reading and
writing of BAM files. katcher and its associated utilities allow for efficient
retrieval and annotation of mapped reads containing any \emph{k}-mer from a set
of interest.

@@

We used contiguous regions of significant associations, which we called
``signals'', as a basis to compare our results with those obtained by previous
studies and to query the genome for candidate genes. We defined the boundaries
of those signals by grouping any significant markers or \emph{k}-mers that
occurred within 250 kb of one another.  We did consider the length of deletions
or \emph{k}-mers in defining the distance between significant associations.

For the GWAS analyses computed from Platypus genotype calls, we retrieved all
markers that had initially been pruned in a region ranging from 50 kb
downstream to 50 kb upstream of significant signals and computed their
\emph{p}-values by running GAPIT using the same parameters as noted above. The
\emph{p}-values thus computed on markers that were initially left out were not
used for initial signal discovery, but provided finer resolution when analyzing
the results of the SNP-based GWAS analyses. Indeed, some markers that had not
been included in the original GWAS may be the causal variant or simply yield
more significant \emph{p}-values than those initially used.

@@

All additional files can be accessed on figshare \citep{kmer-gwas-figshare}.

>>>>

Genome-wide association studies (GWAS) are powerful statistical methods that
detect associations between genotype and phenotype at genome scale. Despite
their power, GWAS frequently fail to pinpoint the causal variant or the gene
controlling a trait at a given locus in crop species. Assessing genetic
variants beyond single-nucleotide polymorphisms (SNPs) could alleviate this
problem, for example by including structural variants (SVs).  In this study, we
tested the potential of SV- and \emph{k}-mer-based GWAS in soybean by applying
these methods to 13 traits.  We also performed conventional GWAS analysis based
on SNPs and small indels for comparison.  We assessed the performance of each
GWAS approach based on results at loci for which the causal genes or variants
were known from previous genetic studies. We found that \emph{k}-mer-based GWAS
was the most versatile approach and the best at pinpointing causal variants or
candidate genes based on the most significantly associated \emph{k}-mers.
Moreover, \textit{k}-mer-based analyses identified promising candidate genes
for loci related to pod color, pubescence form, and resistance to the oomycete
\textit{Phytophthora sojae}. In our dataset, SV-based GWAS did not add value
compared to \textit{k}-mer-based GWAS and may not be worth the time and
computational resources required to genotype SVs at population scale.  Despite
promising results, significant challenges remain regarding the downstream
analysis of \emph{k}-mer-based GWAS.  Notably, better methods are needed to
associate significant \emph{k}-mers with sequence variation. Together, our
results suggest that coupling \emph{k}-mer- and SNP/indel-based GWAS is a
powerful approach for discovering candidate genes in crop species.

@@

The inclusion of structural variants (SVs) as genotype data in GWAS may partly overcome
this issue by providing a set of variants with a potentially high functional impact.
SVs include any variant involving a difference of at least 50 nucleotides, such as
deletions, insertions, or inversions \citep{ho2020}. Such variants are known to
have large phenotypic impacts by disrupting coding or regulatory sequences
\citep{marroni2014}, and have been identified as causal variants in traits such
as resistance to soybean cyst nematode \citep{cook2012}, aluminum tolerance in
wheat \citep{maron2013}, and branching in maize \citep{studer2011}.  SVs have
already been used in GWAS with promising results
\citep[e.g.][]{zhang2015, akakpo2020, dominguez2020, liu2020-pangenome}, but
their use in GWAS is the exception rather than the norm.

Despite their phenotypic impact, SVs are still largely understudied compared to
SNPs because discovering and genotyping them with accuracy is difficult.
Comprehensive detection of SVs requires a combination of several methods that
often need to be tailored to the study species and data available
\citep{alkan2011, ho2020}. Even then, assessments of the performance of SV
discovery and genotyping generally reveal subpar sensitivity and precision
\citep[e.g.][]{cameron2019, chaisson2019, kosugi2019}. Moreover, despite recent
improvements in computational approaches for genotyping SVs
\citep[e.g.][]{siren2021, ebler2022}, such approaches remain relatively
inaccurate compared to SNP genotyping workflows.

Alternative approaches that allow one to assess a large spectrum
of variants in GWAS without genotyping SVs at population scale could be useful.
\cite{voichek2020} developed a \emph{k}-mer-based GWAS approach that may
represent such an alternative method. This approach relies on the presence or
absence of \emph{k}-mers observed in sequence reads in place of variant genotypes for use in
GWAS. Once a presence/absence table of \emph{k}-mers has been generated for the
population under study, this table can be used in association analyses
similarly to SNP or SV genotype calls. One advantage of using \emph{k}-mers is
that they can act as a molecular signature for any type of variant as long as
these variants result in presence/absence of \emph{k}-mers of a given length.
Another advantage is that \emph{k}-mers are not tied to a specific genomic
location and can thus be used to query associations between genotype and
phenotype at genomic locations that are not included in a reference genome.  A
few studies have applied this \emph{k}-mer-based GWAS approach
\citep[e.g.][]{tripodi2021, colque2021} and other similar approaches have been
developed \citep{rahman2018, he2021}, but similarly to the use of SVs in GWAS,
such approaches have not been widely adopted yet.

In this study, we assessed the potential of SV- and \emph{k}-mer-based GWAS in
a major crop, soybean (\emph{Glycine max}). We also conducted conventional GWAS
based on SNPs and indels. In order to test these methods, we analyzed ten
qualitative traits \citep{bandillo2017} and two quantitative traits
\citep{bandillo2015} that have been previously studied using a SNP array
developed for soybean \citep{song2013}. In addition, we analyzed a quantitative
trait (horizontal resistance to \emph{Phytophthora sojae}) that has been
previously studied using whole-genome sequencing (WGS) data
\citep{deronne2022}.  Since the underlying genes and causal variants are
already known for several of these traits, they provide a good test case for
the performance of GWAS methods.  The objectives of our study were to:

@@

We analyzed a total of 13 traits in a population of 363 resequenced \textit{G.
max} accessions using three GWAS approaches that differed in the genotypes
used. The first approach used SNP and indel genotypes called by Platypus.
Despite both SNPs and indels being used, we will refer to this approach as
SNP-based for simplicity. Our SV-based workflow used various sources of data
for SV discovery and a single program (Paragraph) for genotyping SVs at
population scale from Illumina WGS data. Finally, the third GWAS approach
relied on the presence/absence of \textit{k}-mers in the WGS data as genotypes.
Because of the large number of \textit{k}-mers analyzed, we only report results
for those that were considered significantly associated with traits of
interest.

@@

% Pubescence color - T locus
\cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
\textit{T} locus for pubescence color and found a 1-bp deletion as a putative
causal variant at this locus. For all three GWAS approaches, we found a signal
overlapping this gene (Additional file 1: Figure \ref{pubescence_color_all_manhattan}). In
particular, the regions defined by the 1\% most associated variants
identified by the SNP- and \textit{k}-mer-based approaches included
Glyma.06g202300 (Additional file 1: Figure \ref{pubescence_color_all_T_signal}). While the most
significant \textit{k}-mers did occur within the sequence of this gene, they
did not correspond to the documented causal variant (Additional file 1: Figure
\ref{pubescence_color_all_T_gene}d).  Still, the fourth and fifth most strongly
associated \textit{k}-mers corresponded to this causal variant (Additional file 1: Figures
\ref{pubescence_color_all_T_gene}d and \ref{pubescence_color_all_T_kmers}). We
observed that not all accessions with gray pubescence harbored the deleted
nucleotide (Additional file 1: Figure \ref{pubescence_color_all_T_kmers}c), which could indicate
either that other causal variants are resulting in a non-functional F3’H in
this population, or that the causal variant lies elsewhere. As for SNP-based GWAS,
the most significant variant within the sequence of Glyma.06g202300 was the
causal variant, however it was not the most significantly associated variant
overall (Additional file 1: Figures \ref{pubescence_color_all_T_signal}b and
\ref{pubescence_color_all_T_gene}b). SV-based GWAS did not detect the causal variant
at this locus, as expected from its size.

@@

% Pubescence color - Td locus
We conducted a separate GWAS analysis contrasting only accessions with tawny
and light tawny pubescence because the \textit{Td} locus is known to control
these differences in color. Accordingly, we detected signals in this region
using all three approaches, but only SNP- and \textit{k}-mer-based GWAS
identified a signal that overlapped Glyma.03g258700 (Figure
\ref{pubescence-color-main-figure}a-b, and Additional file 1: Figures \ref{pubescence_color_nogray_manhattan} and
\ref{pubescence_color_nogray_Td_signal}).  While the most significant
\textit{k}-mer identified the causal variant documented at this locus
\citep{yan2020}, it was the fourth most significantly associated variant for
the SNP-based approach (Figure \ref{pubescence-color-main-figure}b-d, Additional file 1: Figure 
\ref{pubescence_color_nogray_Td_gene}).  Interestingly, some of the accessions
bearing the haplotype associated with tawny pubescence displayed light tawny
pubescence, suggesting that there may be more than one causal variant at this
locus (Figure \ref{pubescence-color-main-figure}e).

@@

The difference in size between the CNV reported by \cite{liu2020ps} and our
results may be due to the presence of some unresolved sequence (a stretch of
4,635 'N' nucleotides) over that interval in the reference sequence that we used.
Because our SV filtering process removed any SV overlapping 'N' nucleotides in
the reference, we reasoned that the CNV may have been found by our pipeline but
filtered out later on. Indeed, we found a 42.8-kb duplication corresponding
to the one described above among the genotype calls made by smoove. An association
analysis of that single variant using GAPIT found a \textit{p}-value of 1.9e-23,
which would have made it the most significant SV. In this particular case, the
failure of the SV-based GWAS analysis to find identify the causal variant was due
to our filtering parameters and not to limitations of the SV genotype calls per se.

@@

\begin{figure}
	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
	\textit{Pa1} locus for pubescence form. (a) Genome-wide Manhattan plots.
	Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{Pa1} and \textit{Pa2} loci. (b) Zoomed-in Manhattan plots at the \textit{Pa1} locus.
	The vertical dotted lines indicate the location of the
	Glyma.12g213900 gene suggested as a candidate for this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of SNPs, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.12g213900. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrows indicate the
	direction of transcription.
	(d) Manhattan plots at the location of the Glyma.12g213900 gene.
	Vertical lines indicate the location of two non-synonymous
	SNPs associated with appressed pubescence, which may represent causal variants
	at this locus. Panels (c) and (d) share the same x-axis.}
\end{figure}

@@

In conclusion, we used SV-based and \emph{k}-mer-based GWAS to study ten
qualitative traits and three quantitative traits in soybean and
compared the results to conventional SNP-based GWAS. While \emph{k}-mer-based
GWAS proved to be a powerful approach in pinpointing the causal variation or
genes associated with known loci, it is unclear whether conducting SV-based
GWAS is worth the investment. In addition, we used the results from the
\emph{k}-mer- and SNP-based GWAS to suggest candidate genes for a few classical
loci that have yet to be cloned. Based on our results, we believe that an
optimal workflow may involve conducting SNP-based and \emph{k}-mer-based GWAS
in parallel to identify significant signals and candidate genes, potentially in
combination with SV datasets. Given the large number of samples for which WGS
data is now available in several species, applying \emph{k}-mer-based GWAS to
leverage already existing sequencing and phenotypic data appears feasible and
promising.  As part of our work, we developed several computational tools that
should help other researchers with the downstream analysis of
\emph{k}-mer-based GWAS. However, much work remains to be done in
developing state-of-the-art methods for the downstream analysis of significant
\emph{k}-mers. In particular, better approaches are needed for linking
\emph{k}-mers to sequence variation and moving from presence/absence-based
methods to count-based methods.

@@

The SVs called using each of these approaches were merged using SVmerge
\citep{wong2010} and the resulting set of candidate SVs was used for
genotyping. We used Paragraph \citep{chen2019} to genotype the candidate SVs
from the mapped Illumina reads for all accessions. We filtered the genotyped
SVs by setting the genotype calls made from less than two reads to missing and
by removing variants with a minor allele frequency < 0.02 or a proportion of
missing data > 0.5. The resulting dataset of 186,306 genotyped SVs was used for
GWAS. Detailed methods regarding the SV discovery and genotyping steps can be
found in Additional file 1.

@@

\begin{enumerate}
	\item \texttt{FILTER} field set to \texttt{PASS}
	\item minor allele frequency (MAF) $\geq$ 0.02
	\item fraction of missing genotypes $\leq$ 0.5
	\item heterozygosity rate $\leq$ 0.2
\end{enumerate}

@@

Because of clear discrepancies between the observed haplotypes and the reported
phenotypes for some simple traits such as flower color and seed coat color, we
suspected possible errors in the identity of some of the accessions. To
investigate this, we compared the SNP genotypes derived from the SoySNP50K
array and our WGS data at over 32,000 SNPs, and identified 24 accessions with <
90\% concordance between the two genotype datasets (see Additional file 1 for
detailed methods). These samples were removed from the dataset in addition to another
accession that exhibited an atypical GC content suggesting contamination.
Following these filtering steps and the removal of HN019, 363 accessions
remained for GWAS analyses. The concordance between WGS data and SoySNP50K
genotypes is included in Additional file 3 and averaged 98.3\% among retained
lines (Additional file 1: Figure \ref{concordance-histogram}). The phenotypic data used in this study can be found in Additional file 4.

@@

\emph{k}-mers are not intrinsically associated with a particular genomic region
and a tailored analysis was therefore required to associate them with genomic
coordinates. To do so, we queried the mapped reads (in BAM format) of all accessions for matches
to significant \emph{k}-mers for a given trait. Because this is a
computationally demanding operation (there were typically tens of thousands of
significant \emph{k}-mers per trait), we developed a C program called katcher
(\url{https://github.com/malemay/katcher}) that
uses the htslib library \citep{bonfield2021} for efficient reading and
writing of BAM files. katcher and its associated utilities allow for efficient
retrieval and annotation of mapped reads containing any \emph{k}-mer from a set
of interest.

@@

We used contiguous regions of significant variants, which we called
``signals'', as a basis to compare our results with those obtained by previous
studies and to query the genome for candidate genes. We defined the boundaries
of those signals by grouping any significant markers or \emph{k}-mers that
occurred within 250 kb of one another.  We did consider the length of deletions
or \emph{k}-mers in defining the distance between significant variants.

For the GWAS analyses computed from SNP/indel genotype calls, we retrieved all
markers that had initially been pruned in a region ranging from 50 kb
downstream to 50 kb upstream of significant signals and computed their
\emph{p}-values by running GAPIT using the same parameters as noted above. The
\emph{p}-values thus computed on markers that were initially left out were not
used for initial signal discovery, but provided finer resolution when analyzing
the results of the SNP-based GWAS analyses. Indeed, some markers that had not
been included in the original GWAS may be the causal variant or simply yield
more significant \emph{p}-values than those initially used.

@@

### 2023-03-03 additional_files/manuscript.tex
Make a few changes in the text and figures upon proofreading the manuscript

% Defining a new command for signal plot figures
\newcommand{\signalplot}[4]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_signal}
	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zoomed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using three genotype datasets:
	(b) Platypus (SNPs and indels), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
	Panel (a) shows gene models over the genomic interval.
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	#4
	Gray shaded rectangles indicate the region delimited by the top 5\% (Paragraph) or top 1\% (Platypus and \emph{k}-mers) associations in the signal region.
	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	}
	\label{#1_signal}
\end{cfigure}

@@

% Defining a new command for gene plot figures
\newcommand{\geneplot}[5]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_gene}
	\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using three genotype datasets:
	(b) Platypus (SNPs and indels), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
	#5
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	Panel (a) shows gene models over the plotting interval.
	Exons are represented by rectangles whereas introns are represented by horizontal lines.
	Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
	}
	\label{#1_gene}
\end{cfigure}

>>>>

% Defining a new command for signal plot figures
\newcommand{\signalplot}[4]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_signal}
	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zoomed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using three genotype datasets:
	(b) Platypus (SNPs and indels), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
	Panel (a) shows gene models over the genomic interval.
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	#4
	Gray shaded rectangles indicate the region delimited by the top 5\% (SVs) or top 1\% (SNPs/indels and \emph{k}-mers) associations in the signal region.
	In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	}
	\label{#1_signal}
\end{cfigure}

@@

% Defining a new command for gene plot figures
\newcommand{\geneplot}[5]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_gene}
	\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using three genotype datasets:
	(b) Platypus (SNPs and indels), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
	#5
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	In the case of SNPs/indels, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	Panel (a) shows gene models over the plotting interval.
	Exons are represented by rectangles whereas introns are represented by horizontal lines.
	Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
	}
	\label{#1_gene}
\end{cfigure}
